The Astropy Project: Building an open-science project and status of the v2.0 core package

### 2018-02-25 response.txt

We thank the referee for the close reading of the paper and the insightful 
comments.  We provide a response to specific comments from the referee
and the statistical editor below.

@@

We thank the referee for bringing up this important point. After some discussion
within the Astropy community, it became clear that a section regarding inclusion 
is beyond the scope of the current paper. Quantitative analysis of the Astropy 
community demographics is warranted. However, this is a serious undertaking 
beyond the capabilities of our community, and it may require IRB approval. Based 
on the referee's comment, we have established initial contact with the NumFOCUS 
Diversity & Inclusion in Scientific Computing (DISC) program. Results will not 
be available in time for this paper. We have opted to remove "inclusive" from 
the paper title to make room for future study and dialogue.

> The phase 'open development' is used a few times but isn't really defined or 
> connected to the broader open source ecosystem.  What other communities are 
> working in this way? How novel is 'open development' by their assessment? For 
> example, other communities have adopted policies that encourage community 
> contribution - most notable is the Node.js Foundation's contribution policies
> designed to promote 'Healthy Open Source' 
> https://medium.com/the-node-js-collection/healthy-open-source-967fa8be7951 . 
> How does the Astropy community efforts compare to this?

This question is at the intersection of two communities. On the one
side, there is the open source community, which by itself consists of
many different projects with vastly different norms and
processes. Many terms, "open develoment" being one of them, are used
with slightly different meanings in different parts of this
community. The model of astropy - the development happens openly on
github and anyone with an internet connection can comment on issues or
propose a pull request, but only a select set of maintainers can merge
them - is by no means special. Many open source communities work in a
very similar way, although they may use different terminology to
describe it. On the other hand, there is the astrophysics community,
and while there are several open source and open development (in the
way astropy uses the term) projects in astronomy, astropy is probably
the biggest one. Many astronmers are not involved in any open
development (also many make use of open source projects they are not
involved in e.g. GNU/Linux), and the concept is new to many of
them. This article caters mostly to the astrophysics community, so we
decided to add a definition, but we refrain from a comparative
discussion with other open source projects.

@@

________________________

> Note from statistical editor.

> In sec 3.11.3, a sentence can be added warning users about difficulties in 
> interpreting the statistical significance of Lomb -Scargle periodogram peaks. 
> This is important because unreliable interpretation is common, and warnings 
> are not given elsewhere (e.g. NASA Exoplanet Archive Periodogram Service). 
> A good reference is Vanderplas 2017 (https://arxiv.org/abs/1703.09824) 
> soon to appear in the ApJ Suppl.

A short note directing the reader to Vanderplas (2007) was included.

### 2018-02-25 main.tex
explain open development
Closes #140

The \astropy project aims to provide an open-source and open-development core
package (\astropypkg) and an ecosystem of \emph{affiliated packages} that
support astronomical functionality in the \python programming language.
The \astropypkg core package is now a feature-rich library of sufficiently
general tools and classes that supports the development of more specialized
code. An example of such functionality is reading and writing FITS files: It would be
time consuming and impractical for multiple groups to implement the FITS
standard \citep{FITS} and maintain software for such a general-purpose need.
Another example of such a common task is in dealing with representations of and
transformations between astronomical coordinate systems.

@@

\subsection{Astropy development model}

Code is contributed to the \astropypkg core package or modified through ``pull
requests'' (via \github\footnote{\url{https://github.com/astropy/astropy/}})
that often contain several \texttt{git} commits.
Pull requests may fix bugs, implement new features, or improve or modify the
infrastructure that supports the development and maintenance of the package.
Individual pull requests are generally limited to a single conceptual addition
or modification to make code review tractable.
Pull requests that modify or add code to a specific subpackage must be reviewed
and approved by one of the subpackage maintainers before they are merged into
the core codebase.
Bugs and feature requests are reported via the \github issue tracker and labeled
with a set of possible labels that help classify and organize the issues.
The development workflow is detailed in the \astropypkg
documentation.\footnote{\emph{How to make a code contribution},
\url{http://docs.astropy.org/en/latest/development/workflow/development_workflow.html}}

\begin{figure}
\includegraphics[width=\textwidth]{commits.pdf}
\caption{%
    \emph{Left panel}: Distribution of number of commits per committer.
    \emph{Middle panel}: Cumulative number of commits to the \astropypkg core
    package over time.
    \emph{Right panel}: Number of unique committers per month to the
    \astropypkg core package.
    \label{fig:ncommits}
}
\end{figure}

@@

\subsection{APEs - Astropy Proposals for Enhancement}

Central to the success of \astropy is an open environment where anybody can
contribute to the project.
This model leads to an ``organic'' growth, where features are
implemented by different people with different programming styles and
interfaces.
Thus, \astropy has a mechanism to more formally propose significant changes to
the core package (e.g., re-writing the coordinates subpackage; \citealt{ape5}),
to plan out major new features (e.g., a new file format; \citealt{ape6}), or
institute new organization-wide policies (e.g., adopting a code of conduct;
\citealt{ape8}).
This mechanism is called ``Astropy Proposal for Enhancement'' (APE) and is
modeled after the ``Python Enhancement Proposals'' (PEP) that guide the
development of the \python programming language.
In an APE, one or more authors describe in detail the proposed changes or
additions, including a rationale for the changes, how these changes will be
implemented, and in the case of code, what the interface will be \citep{ape1}.
The APEs are discussed and refined by the community before much work is invested
into a detailed implementation; anyone is welcome to contribute to these
discussions during the open consideration period. APEs are proposed via pull
requests on a dedicated GitHub repository\footnote{\url{https://github.com/astropy/astropy-APEs}};
anyone can therefore read the proposed APEs and leave in-line comments.
When a community consensus emerges, the APEs are accepted and become
the basis for future work.
In cases where consensus cannot be reached, the
\astropy coordination committee may decide to close the discussion and
make an executive decision based on the community input on the APE in question.

>>>>

The \astropy project aims to provide an open-source and open-development core
package (\astropypkg) and an ecosystem of \emph{affiliated packages} that
support astronomical functionality in the \python programming language.
``Open-development'' describes a process where anybody with an internet connection can suggest changes to the source code and contribute his or her opinion when new features, bug fixes, or other code changes are discussed (see Sect.~\ref{sect:devmod} and \ref{sect:ape} of how this is organized in practice).
The \astropypkg core package is now a feature-rich library of sufficiently
general tools and classes that supports the development of more specialized
code. An example of such functionality is reading and writing FITS files: It would be
time consuming and impractical for multiple groups to implement the FITS
standard \citep{FITS} and maintain software for such a general-purpose need.
Another example of such a common task is in dealing with representations of and
transformations between astronomical coordinate systems.

@@

\subsection{Astropy development model}
\label{sect:devmod}
Code is contributed to the \astropypkg core package or modified through ``pull
requests'' (via \github\footnote{\url{https://github.com/astropy/astropy/}})
that often contain several \texttt{git} commits.
Pull requests may fix bugs, implement new features, or improve or modify the
infrastructure that supports the development and maintenance of the package.
Individual pull requests are generally limited to a single conceptual addition
or modification to make code review tractable.
Pull requests that modify or add code to a specific subpackage must be reviewed
and approved by one of the subpackage maintainers before they are merged into
the core codebase.
Bugs and feature requests are reported via the \github issue tracker and labeled
with a set of possible labels that help classify and organize the issues.
The development workflow is detailed in the \astropypkg
documentation.\footnote{\emph{How to make a code contribution},
\url{http://docs.astropy.org/en/latest/development/workflow/development_workflow.html}}

\begin{figure}
\includegraphics[width=\textwidth]{ncommits.pdf}
\caption{%
    \emph{Left panel}: Distribution of number of commits per committer.
    \emph{Right panel}: Cumulative number of commits to the \astropypkg core
    package over time.
    \label{fig:ncommits}
}
\end{figure}

@@

\subsection{APEs - Astropy Proposals for Enhancement}
\label{sect:ape}
Central to the success of \astropy is an open environment where anybody can
contribute to the project.
This model leads to an ``organic'' growth, where features are
implemented by different people with different programming styles and
interfaces.
Thus, \astropy has a mechanism to more formally propose significant changes to
the core package (e.g., re-writing the coordinates subpackage; \citealt{ape5}),
to plan out major new features (e.g., a new file format; \citealt{ape6}), or
institute new organization-wide policies (e.g., adopting a code of conduct;
\citealt{ape8}).
This mechanism is called ``Astropy Proposal for Enhancement'' (APE) and is
modeled after the ``Python Enhancement Proposals'' (PEP) that guide the
development of the \python programming language.
In an APE, one or more authors describe in detail the proposed changes or
additions, including a rationale for the changes, how these changes will be
implemented, and in the case of code, what the interface will be \citep{ape1}.
The APEs are discussed and refined by the community before much work is invested
into a detailed implementation; anyone is welcome to contribute to these
discussions during the open consideration period. APEs are proposed via pull
requests on a dedicated GitHub repository\footnote{\url{https://github.com/astropy/astropy-APEs}};
anyone can therefore read the proposed APEs and leave in-line comments.
When a community consensus emerges, the APEs are accepted and become
the basis for future work.
In cases where consensus cannot be reached, the
\astropy coordination committee may decide to close the discussion and
make an executive decision based on the community input on the APE in question.
